---
title: "Mario Kart Analysis"
execute:
  echo: true
  warning: false
  message: false
  error: true
format: 
  html:
    self-contained: true
    slide-number: true
    footer: "STA6349 - Applied Bayesian Analysis - Fall 2025"
    width: 1600
    height: 900
    df-print: paged
    html-math-method: katex
    code-fold: false
    code-tools: false
    incremental: false
editor: source
---

## Example: Mario Kart

- In Mario Kart 8 Deluxe, item boxes give different items depending on race position. To reduce the "position bias," only item boxes opened while the racer was in mid-pack (positions 4–10) were recorded. 

- You want to estimate the probability that an item box yields a Red Shell. When playing the Special Cup, only 31 red shells were seen in 114 boxes opened by mid-pack racers.

- Find the posterior distribution under two priors:
    1. Flat/uninformative prior, Beta(1,1).
    2. Beta($\alpha$, $\beta$) of your choosing.

*Find the posterior under the flat prior, Beta(1,1).*

```{r}
#| echo: false
library(tidyverse)
library(bayesrules)
```

```{r, fig.align="center"}
plot_beta_binomial(alpha = 1, beta = 1,
                   y = 31, n = 114) +
  theme_bw()
```

*Find the posterior under a prior of your choosing.*

```{r, fig.align="center"}
plot_beta_binomial(alpha = 8, beta = 4,
                   y = 31, n = 114) +
  theme_bw()

plot_beta_binomial(alpha = 23, beta = 57,
                   y = 31, n = 114) +
  theme_bw()
```


- Suppose that we know the individual data points for the Special Cup:
    - Cloudtop Cruise, of 28 boxes, 6 had a red shell.
    - Bone-Dry Dunes, of 27 boxes, 8 had a red shell.
    - Bowser's Castle, of 29 boxes, 12 had a red shell.
    - Rainbow Road, of 30 boxes, 5 had a red shell.

- Prove sequentiality to yourself.

    - Analyze the data race-by-race in this order: Cloudtop Cruise, Bone-Dry Dunes, Bowser's Castle, and Rainbow Road.
    
    - Analyze the data race-by-race in this order: Rainbow Road, Bowser's Castle, Cloudtop Cruise, and Bone-Dry Dunes.
    
*Analyze the data race-by-race in this order: Cloudtop Cruise, Bone-Dry Dunes, Bowser's Castle, and Rainbow Road.*

```{r}
summarize_beta_binomial(alpha = 1, beta = 1,
                        y = 6+8+12+5, n = 28+27+29+30) # overall


summarize_beta_binomial(alpha = 1, beta = 1,
                        y = 6, n = 28) # cloudtop cruise
summarize_beta_binomial(alpha = 7, beta = 23,
                        y = 8, n = 27) # bone-dry dunes
summarize_beta_binomial(alpha = 15, beta = 42,
                        y = 12, n = 29) # bowser's castle
summarize_beta_binomial(alpha = 27, beta = 59,
                        y = 5, n = 30) # bowser's castle
```

Final distribution is Beta(32, 84).

```{r}
summarize_beta_binomial(alpha = 27, beta = 59,
                        y = 5, n = 30) # rainbow road
```

Final distribution is Beta(32, 84).

*Analyze the data race-by-race in this order: Rainbow Road, Bowser's Castle, Cloudtop Cruise, and Bone-Dry Dunes.*

In order:

```{r}
summarize_beta_binomial(alpha = 1, beta = 1,
                        y = 5, n = 30) # rainbow road
summarize_beta_binomial(alpha = 6, beta = 26,
                        y = 12, n = 29) # bowser's castle
summarize_beta_binomial(alpha = 18, beta = 43,
                        y = 6, n = 28) # cloudtop cruise
summarize_beta_binomial(alpha = 24, beta = 65,
                        y = 8, n = 27) # bone-dry dunes
```

Final distribution is Beta(32, 84).

All together: 

```{r}
summarize_beta_binomial(alpha = 1, beta = 1,
                        y = 5+12+6+8, n = 30+29+28+27)
```

Final distribution is Beta(32, 84).